---
title: "Homework 3"
author: Laura Gomez
output: github_document
---

```{r setup}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

## Problem 1

How many aisles are there, and which aisles are the most items ordered from?
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


## Problem 2

Import and clean the accelerometer dataset. Then add the weekday vs weekend variable and encode data with reasonable variable classes
```{r accelerometer_tidy_df}
accelerometer_df = read_csv("./Data/accel_data.csv") %>%
  janitor::clean_names() %>%
  drop_na()  %>%
  pivot_longer(activity_1:activity_1440, names_to = 'time', names_prefix = 'activity_', values_to = 'activity_count') %>% 
  mutate(
    day = as.factor(day), 
    time = as.numeric(time)) %>% 
  mutate(
    weekend = ifelse(day == c("Saturday", "Sunday"),1,0), 
    weekend = case_when(weekend == 1 ~ "weekend", weekend == 0 ~ "weekday" )) %>%
  mutate(
    day = forcats::fct_relevel(
      day, "Monday", "Tuesday","Wednesday", "Thursday", "Friday", "Saturday", "Sunday" ))
```

The accelerometer dataset contains information self-reported physical activity over a 24 hour cycle. The following variables exist in the dataset `r ls(accelerometer_df)`. There are a total of `r nrow(accelerometer_df)` rows and `r ncol(accelerometer_df)` columns in our final dataset. 

Aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
```{r total activity_df}
aggregate_activity_df = 
  accelerometer_df %>% 
  group_by(time) %>% 
  summarize(total_activity_df = sum(activity_count, na.rm = TRUE)) 
#%>% knitr::kable(aggregate_activity_df)
```

Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accelerometer_df %>%
  ggplot(aes(x = time, y = activity_count, color = day)) + 
  geom_point() 
#fix order of data 
```
Describe in words any patterns or conclusions you can make based on this graph.


## Problem 3

Import data
```{r}
library(p8105.datasets)
data("ny_noaa")
```
To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue.
```{r}
ny_noaa_df = ny_noaa %>% separate('date', c( 'year','month','day')) %>% 
  mutate(
    year = factor(year), 
    month = factor(month), 
    day = factor(day), 
    prcp =
      as.integer(prcp), 
    tmax = as.integer(tmax), 
    tmin = as.integer(tmin)) %>% 
  drop_na()
```



