---
title: "Homework 3"
author: Laura Gomez
output: github_document
---

```{r setup}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

## Problem 1

How many aisles are there, and which aisles are the most items ordered from?
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


## Problem 2

Import and clean the accelerometer dataset. Then add the weekday vs weekend variable and encode data with reasonable variable classes
```{r accelerometer_tidy_df}
accelerometer_df = read_csv("./Data/accel_data.csv") %>%
  janitor::clean_names() %>%
  drop_na()  %>%
  pivot_longer(activity_1:activity_1440, names_to = 'time', names_prefix = 'activity_', values_to = 'activity_count') %>% 
  mutate(
    day = as.factor(day), 
    time = as.numeric(time)) %>% 
  mutate(
    weekend = ifelse(day == c("Saturday", "Sunday"),1,0), 
    weekend = case_when(weekend == 1 ~ "weekend", weekend == 0 ~ "weekday" ))
```

The accelerometer dataset contains information self-reported physical activity over a 24 hour cycle. The following variables exist in the dataset `r ls(accelerometer_df)`. There are a total of `r nrow(accelerometer_df)` rows and `r ncol(accelerometer_df)` columns in our final dataset. 

